Cominuscule subvarieties of flag varieties

### 2023-10-05 ms.tex
Add files via upload

\thanks{This research was supported in part by the International Centre for Theoretical Sciences (ICTS) during a visit for participating in the program - Analytic and Algebraic Geometry (Code: ICTS/aag2018/03).
This research was completed at the University of Catania, thanks to the hospitality of the university and of Francesco Russo.
Thanks to Indranil Biswas, Anca \texorpdfstring{Musta\c{t}\u{a}}{Mustata} and Andrei \texorpdfstring{Musta\c{t}\u{a}}{Mustata} for help with algebraic geometry.}
\keywords{flag variety, Hermitian symmetric space}
%\subjclass[2000]{Primary 53B21; Secondary 53C56, 53A55}
\date{\today}

@@

@@

Denote the Lie algebras of \(P\subseteq G\) by \(\LieP\subset\LieG\). 
One can select a Cartan subgroup of \(G\) lying inside \(P\), whose positive root spaces all lie in \(\LieP\).
A simple root \(\aa\) is \(P\)-\emph{compact} (\emph{compact} if \(P\) is understood) if the root space of \(-\aa\) belongs to the Lie algebra of \(P\).
Each flag variety is determined uniquely, up to finite central extension of \(G\) and up to isomorphism, by the Dynkin diagram of \(G\) decorated with \(\dynkin{A}{*}\) on each compact simple root and \(\dynkin{A}{x}\) on each noncompact root \cite{Borel:1991} p. 197 Proposition 14.18.
\subsection{Reducible flag varieties}\label{subsec:reducible}
The center of any complex semisimple Lie group \(G\) lies in every maximal torus, so in every Cartan subgroup \cite{Borel:1991} p. 220, so in every parabolic subgroup, so acts trivially on every flag variety.
An \emph{irreducible} flag variety is a flag variety \((X,G)\) with \(G\) a simple Lie group, and with only the center of \(G\) acting trivially.
Every flag variety \((X,G)\), up to finite central extension of \(G\), admits a factorization 
\begin{align*}
X&=X_0 \times X_1 \times X_2 \times \dots \times X_s, \\
G&=G_0 \times G_1 \times G_2 \times \dots \times G_s, 
\end{align*}
into irreducible flag varieties \(\pr{X_i, G_i}\), \(i>0\), and a point \(X_0=\set{x_0}\), unique up to permutation of the \((X_i,G_i)\) for \(i>0\) and isomorphism.
The flag variety \((X,G)\) is effective if and only if all \((X_i,G_i))\) are effective, i.e. if and only if \(G_0\) is trivial and \(G_1,\dots,G_s\) are in adjoint form.
\subsection{Cominuscule varieties}
A flag variety is \emph{cominuscule} if \(\LieG/\LieP=T_{x_0} X\) is a sum of irreducible complex algebraic \(P\)-modules.
This occurs just when there is a compact subgroup \(K\subseteq G\) so that \(\pr{X,K}\) is a compact Hermitian symmetric space \cite{Kostant:1961} p. 379 Proposition 8.2, \cite{Baston/Eastwood:1989} p. 26.
Some authors prefer the term \emph{compact Hermitian symmetric space}, \emph{cominuscule Grassmannian}, or \emph{generalized Grassmannian} to \emph{cominuscule variety}.
Every effective cominuscule variety is a product of the following irreducible effective cominuscule varieties \cite{Kobayashi/Nagano:1964} theorem 1 p. 401:\par\noindent{}%
%%\input{hss}
\begingroup
\small
\begin{longtabl}{@{}%
>{\columncolor[gray]{.9}$}r<{$}%
>{\columncolor[gray]{.93}}p{2.6cm}%
>{\columncolor[gray]{.9}$}r<{$}%
>{\columncolor[gray]{.93}}l@{}}
\toprule
G&\(G/P\)&\operatorname{dim}&\text{description}\\
\midrule
\endfirsthead
\toprule
G&\(G/P\)&\operatorname{dim}&\text{description}\\
\midrule
\endhead
\bottomrule
\endfoot
\bottomrule
\endlastfoot
A_r&\dynkin{A}{**.*x*.**}&k(r+1-k)&Grassmannian of $k$-planes in $\C[r+1]$
\\
B_r&\dynkin[parabolic=1]{B}{}&2r-1&quadric hypersurface in $\Proj{2r}$\\
C_r&\dynkin[parabolic=16]{C}{}&\frac{r(r+1)}{2}&space of Lagrangian $r$-planes in $\C[2r]$\\
D_r&\dynkin[parabolic=1]{D}{}&2r-2&quadric hypersurface in $\Proj{2r-1}$\\
D_r&\dynkin[parabolic=32]{D}{}&\frac{r(r-1)}{2}& %one component of the variety of 
space of null $r$-planes in $\C[2r]$ \\
E_6&\dynkin[parabolic=1]{E}{6}&16&complexified octave projective plane\\
E_7 &\dynkin[parabolic=64]{E}{7}&27&space of null octave \(3\)-planes in octave \(6\)-space
\end{longtabl}
\endgroup
%% end input
\subsection{Structure of linear algebraic groups}
A subgroup of a linear algebraic group is \emph{unipotent} if it consists of unipotent linear maps.
Every complex linear algebraic group \(G\) has a \emph{unipotent radical}, the unique maximal unipotent normal subgroup, which is a closed complex linear algebraic subgroup \cite{Borel:1991} p. 85 Theorem 4.5, p. 86 Theorem 4.7, p. 157 11.21.

A complex linear algebraic group is \emph{reductive} if it contains a Zariski dense compact subgroup \cite{Knapp:2002} p. 382 Proposition 7.15.
Every complex linear algebraic group \(G\) has a \emph{reductive Levi factor}, i.e. a maximally reductive complex linear algebraic subgroup,  unique up to conjugacy  \cite{Knapp:2002} p. 478., so that \(G\) is a semidirect product of its reductive Levi factor and its unipotent radical.
Any Cartan subgroup is therefore a subgroup of the Levi factor, after perhaps a conjugacy.
Every compact subgroup lies in a maximal compact subgroup, which is unique up to conjugacy, so lies in the Levi factor up to conjugacy \cite{Hilgert.Neeb:2012} p. 531 Theorem 14.1.3.
After perhaps extending by some finite group of order a power of \(2\), the Weyl group embeds in \(G\) as a finite subgroup \cite{Tits:1966}, hence compact, so this group lies in the Levi factor up to conjugacy.

\subsection{Structure of flag variety stabilizers}
Take a flag variety \((X,G)\). 
Take the stabilizer \(P\defeq G^{x_0}\) of a point \(x_0\in X\).

@@

Every flag variety \(X=G/P\) has parabolic subgroup \(P:=G^{x_0}\) connected and containing a Borel subgroup \(B\subseteq P\) \cite{Borel:1991} p. 197 Proposition 14.18.

@@

(We can define such an automorphism explicitly as \(e_{\alpha}\mapsto -e_{-\alpha}\) on root vectors in a Chevellay basis; see section~\vref{subsubsection:ChevalleyBases}.)
Our automorphism sends \(P\) to its ``opposite'' \(aP\) parabolic subgroup, so gives \(P\cap aP=\GZ\) \cite{Cap/Slovak:2009} p. 294, \cite{Borel:1991} p. 198 Proposition 14.21

@@

Clearly \(\breve{G}\) has Lie algebra containing all root vectors of all \(P\)-maximal and \(P\)-minimal roots, and is generated by the one parameter subgroups of those root vectors.

@@

So \(\breve{X}=\breve{G}/\breve{P}=G'/P'\) is the Grassmannian of linear subspace of dimension \(n_1\) inside \(V_1\oplus V_4\).

@@

@@

We will see \vpageref{subsection:Freedom} that the associated cominiscule subvariety of an irreducible flag variety \((X,G)\) has strictly maximal dimension symmetry group among subvarieties of \(X\) satisfying an open condition on tangent spaces of the smooth locus.
We will see that all smooth subvarieties satisfying that open condition are homogeneous, and all but the associated cominiscule itself have smaller dimensional automorphism group than the associated cominiscule.

@@

For a flag variety \(X=G/P\), the root system is also \emph{\(P\)-graded} by the same sum, but only over the noncompact simple roots.

@@

To see the heights, for each root on the chosen side of the hyperplane, draw a parallel hyperplane through that root:

@@

A \emph{parabolic subgroup} of a reductive Lie group \(G\) is a subgroup \(P\subseteq G\) so that \(X=G/P\) is a smooth complex projective variety; it follows that \(P\) is a connected complex subgroup  \cite{Borel:1991} p. 154 Theorem 11.16.

@@

Every parabolic subgroup of a connected reductive Lie group is connected.

@@

@@

expanding out into a sum with entirely positive coefficients unless \(g\) is a sum of \(P\)-maximal roots, so the center \(Z=Z_{G_+}\) of \(G_+\) consists precisely of the exponentials of elements of

@@

the sum being over the \(P\)-maximal roots, i.e. \(Z\) is connected.
By definition of \(\breve{G}\), \(\breve{G}:=\left<Z,aZ\right>\) is connected, since \(Z\) is connected. \end{proof}
Two associated cominiscule subvarieties are tangent at a point just when they are equal.
Suppose that \(T_{x_0}\breve{X}=T_{x_0}g\breve{X}\).
We can replace \(g\) by \(gh\) for any element \(h\in\breve{G}\), and \(\breve{G}\) acts transitively on \(\breve{X}\) by definition of \(\breve{X}\), so we can arrange that \(gx_0=x_0\), i.e. write \(g\) as \(p\) and we can assume that \(p\in P\).
These tangent spaces are
\breve\LieG/\breve\LieP=\Ad_p\left(\breve\LieG/\breve\LieP\right)\subseteq\LieG/\LieP.
So
\breve\LieG=\Ad_p\breve\LieG\subseteq\LieG.
Hence taking exponentials, the identity components match:
\breve{G}=\Ad_p\breve{G}.
So their orbits are the same.

@@

@@

It also suffices to assume that \((X,G)\) is an irreducible flag variety, as otherwise it is a product of irreducibles.
The Lie algebra \(\LieZ\) of the center \(Z\) of the unipotent radical of \(P\) is the sum of the root spaces of the \(P\)-maximal roots (i.e. roots of maximal \(P\)-grade).

@@

So \(a\LieZ\) is the sum of the root spaces of the roots of minimal \(P\)-grade.

@@

@@

Since \(\LieZ\) is an irreducible \(\GZ\)-module, if we start at the highest root, say \(\aa_0\), we can pass from it via root strings to get to any \(P\)-maximal root, repeatedly passing between \(P\)-maximal roots \(\aa,\bb\) by subtracting a \(\breve{P}\)-compact positive root \(\aa-\bb\), so that bracketing a root vector \(\XX{\aa-\bb}\) of root \(\aa-\bb\) takes \(\XX{\aa}\) to a nonzero multiple of \(\XX{\bb}\).
Hence \(\LieZ\) is an irreducible \(\breveLieGZ\)-module, hence an irreducible \(\breve\LieP\)-module.
As we have seen~\vpageref{subsec:reducible}, the number of irreducible modules of the parabolic subgroup is the number of irreducible factors of the flag variety, so \(\breve{X}\) is an irreducible flag variety.

We next prove that \((\breve{X},\breve{G})\) is effective.
Since \(\breve{P} \subset \breve{G}\) is a parabolic subgroup, we can apply our automorphism \(a\) to \(\breve{P}\), defined directly on roots of \(G\) and taking \(\breve{G}\)-roots to one another, to swap its \(\breve{P}\)-maximal and \(\breve{P}\)-minimal roots.
Since all roots of \(\breve{G}\) are roots of \(G\), this automorphism applies to \(G\) leaving \(\breve{G}\) invariant, and swaps its \(P\)-maximal and \(P\)-minimal roots.
So \(\GN \cap a \GN=1\) and \(a\GZ=\GZ\).
An element \(g \in \breve{G}\) acts trivially on \(\breve{X}\) just when \(gg_1\breve{P}=g_1\breve{P}\) for all \(g_1 \in \breve{G}\), i.e. just when \(g\) lies in all \(\breve{G}\)-conjugates of \(\breve{P}\).
But then \(g \in \breve{P} \cap a\breve{P} \subset P \cap aP=1\).
\end{proof}
\subsection{Automorphisms of the associated cominiscule}
\begin{example}
The flag variety \((X,G)=(B_2/P,B_2)\), where \(P\subseteq B_2\) is the Borel subgroup:
\[
\drawroots{B}{3} 
\]
has associated cominiscule
\[
\begin{tikzpicture}[baseline=-.5]
\begin{rootSystem}{B}
\roots
\parabolic{3}
\draw[/root system/grading,line width=.3cm] (square cs:x=1,y=1) -- (square cs:x=-1,y=-1);
\end{rootSystem}
\end{tikzpicture}
\]
a rational curve \((\breve{X},\breve{G})=(\Proj{1},\PSL{2})\), invariant under not only its automorphism group \(\breve{G}=A_1=\PSL{2}\), which we see in our diagram, but also, as we will see, under rescaling by this root:
\[
\begin{tikzpicture}[baseline=-.5]
\begin{rootSystem}{B}
\roots
\parabolic{3}
\draw[/root system/grading,line width=.3cm] (square cs:x=1,y=1) -- (square cs:x=-1,y=-1);
%\draw (square cs:x=-1,y=1) circle (2pt);
\draw[latex-,shorten <=1mm] (square cs:x=1,y=-1) --++ (.3cm,-.3cm);
\end{rootSystem}
\end{tikzpicture}
\]
The root is perpendicular to the \(\breve{G}\) root system, so commutes with the \(\breve{G}\) root vectors.
The vector field on \(X\) associated to the root vector of that root vanishes at our chosen point \(x_0\in X\) stabilized by \(P\), since the root lies in the Lie algebra of \(P\).
Since the vector field commutes with those of \(\breve{G}\), it is \(\breve{G}\)-invariant.
Since \(\breve{G}\) acts transitively on \(\breve{X}\), our vector field vanishes at all points of \(\breve{X}\).
Hence, in this example, the automorphism group \(G_{\breve{X}}\) of \(\breve{X}\) as a subvariety of \(X\) is larger than the automorphism group \(\breve{G}\) as a flag variety, containing the exponential of this root.

>>>>

\thanks{This research was supported in part by the International Centre for Theoretical Sciences (ICTS) during a visit for participating in the program - Analytic and Algebraic Geometry (Code: ICTS/aag2018/03).
This research was largely written at the University of Catania, thanks to the hospitality of the university and of Francesco Russo.
Thanks to Indranil Biswas, Anca \texorpdfstring{Musta\c{t}\u{a}}{Mustata} and Andrei \texorpdfstring{Musta\c{t}\u{a}}{Mustata} for help with algebraic geometry.
This article/publication is based upon work from COST Action CaLISTA, CA21109, supported by COST (European Cooperation in Science and Technology) \url{www.cost.eu}.}
\keywords{flag variety, Hermitian symmetric space}
%\subjclass[2000]{Primary 53B21; Secondary 53C56, 53A55}
\date{\today}

@@

A Lie group \(G\) acts \emph{almost effectively} on a manifold \(X\) if the elements of \(G\) fixing every point of \(X\) form a finite subgroup. 

@@

Denote the Lie algebras of \(P\subseteq G\) by \(\LieP\subset\LieG\). 
One can select a Cartan subgroup of \(G\) lying inside \(P\), whose positive root spaces all lie in \(\LieP\).
A simple root \(\aa\) is \(P\)-\emph{compact} (\emph{compact} if \(P\) is understood) if the root space of \(-\aa\) belongs to the Lie algebra of \(P\).
Each flag variety is determined uniquely, up to finite central extension of \(G\) and up to isomorphism, by the Dynkin diagram of \(G\) decorated with \(\dynkin{A}{*}\) on each compact simple root and \(\dynkin{A}{x}\) on each noncompact root \cite{Borel:1991} p. 197 Proposition 14.18.
\subsection{Reducible flag varieties}\label{subsec:reducible}
The center of any complex semisimple Lie group \(G\) lies in every maximal torus, so in every Cartan subgroup \cite{Borel:1991} p. 220, so in every parabolic subgroup, so acts trivially on every flag variety.
An \emph{irreducible} flag variety is a flag variety \((X,G)\) with \(G\) a simple Lie group, and with only the center of \(G\) acting trivially.
Every flag variety \((X,G)\), up to finite central extension of \(G\), admits a factorization 
\begin{align*}
X&=X_0 \times X_1 \times X_2 \times \dots \times X_s, \\
G&=G_0 \times G_1 \times G_2 \times \dots \times G_s, 
\end{align*}
into irreducible flag varieties \(\pr{X_i, G_i}\), \(i>0\), and a point \(X_0=\set{x_0}\), unique up to permutation of the \((X_i,G_i)\) for \(i>0\) and isomorphism.
The flag variety \((X,G)\) is effective if and only if all \((X_i,G_i)\) are effective, i.e. if and only if \(G_0=\set{1}\) is trivial and \(G_1,\dots,G_s\) are in adjoint form.
\subsection{Cominuscule varieties}
A flag variety is \emph{cominuscule} if \(\LieG/\LieP=T_{x_0} X\) is a sum of irreducible complex algebraic \(P\)-modules.
This occurs just when there is a compact subgroup \(K\subseteq G\) so that \(\pr{X,K}\) is a compact Hermitian symmetric space \cite{Kostant:1961} p. 379 Proposition 8.2, \cite{Baston/Eastwood:1989} p. 26.
Some authors prefer the term \emph{compact Hermitian symmetric space}, \emph{cominuscule Grassmannian}, or \emph{generalized Grassmannian} to \emph{cominuscule variety}.
Every effective cominuscule variety is a product of the following irreducible effective cominuscule varieties \cite{Kobayashi/Nagano:1964} theorem 1 p. 401:\par\noindent{}%
%%\input{hss}
\begingroup
\small
\begin{longtabl}{@{}%
>{\columncolor[gray]{.9}$}r<{$}%
>{\columncolor[gray]{.93}}p{2.6cm}%
>{\columncolor[gray]{.9}$}r<{$}%
>{\columncolor[gray]{.93}}l@{}}
\toprule
G&\(G/P\)&\operatorname{dim}&\text{description}\\
\midrule
\endfirsthead
\toprule
G&\(G/P\)&\operatorname{dim}&\text{description}\\
\midrule
\endhead
\bottomrule
\endfoot
\bottomrule
\endlastfoot
A_r&\dynkin[labels={,,,k,,,}]{A}{**.*x*.**}&k(r+1-k)&Grassmannian of $k$-planes in $\C[r+1]$
\\
B_r&\dynkin[parabolic=1]{B}{}&2r-1&quadric hypersurface in $\Proj{2r}$\\
C_r&\dynkin[parabolic=16]{C}{}&\frac{r(r+1)}{2}&space of Lagrangian $r$-planes in $\C[2r]$\\
D_r&\dynkin[parabolic=1]{D}{}&2r-2&quadric hypersurface in $\Proj{2r-1}$\\
D_r&\dynkin[parabolic=32]{D}{}&\frac{r(r-1)}{2}& %one component of the variety of 
space of null $r$-planes in $\C[2r]$ \\
E_6&\dynkin[parabolic=1]{E}{6}&16&complexified octave projective plane\\
E_7 &\dynkin[parabolic=64]{E}{7}&27&space of null octave \(3\)-planes in octave \(6\)-space
\end{longtabl}
\endgroup
%% end input
\subsection{Structure of linear algebraic groups}
A complex linear map is \emph{unipotent} if its only eigenvalue is \(1\).
A subgroup of a linear algebraic group is \emph{unipotent} if it consists of unipotent linear maps.
Every complex linear algebraic group \(G\) has a \emph{unipotent radical}, the unique maximal unipotent normal subgroup, which is a closed complex linear algebraic subgroup \cite{Borel:1991} p. 85 Theorem 4.5, p. 86 Theorem 4.7, p. 157 11.21.

A complex linear algebraic group is \emph{reductive} if it contains a Zariski dense compact subgroup \cite{Knapp:2002} p. 382 Proposition 7.15.
Every complex linear algebraic group \(G\) has a \emph{reductive Levi factor}, i.e. a maximally reductive complex linear algebraic subgroup,  unique up to conjugacy, so that \(G\) is a semidirect product of its reductive Levi factor and its unipotent radical \cite{Borel:1991} p. 158 11.22.
Any Cartan subgroup is therefore a subgroup of the Levi factor, after perhaps a conjugacy.
Every compact subgroup lies in a maximal compact subgroup, which is unique up to conjugacy, so lies in the Levi factor up to conjugacy \cite{Hilgert.Neeb:2012} p. 531 Theorem 14.1.3.
After perhaps extending by some finite group of order a power of \(2\), the Weyl group embeds in \(G\) as a finite subgroup \cite{Tits:1966}, hence compact, so this group lies in the Levi factor up to conjugacy.

\subsection{Parabolic subgroups}
A Zariski closed subgroup \(P\subseteq G\) of a connected linear algebraic group \(G\) is \emph{parabolic} if \(X:=G/P\) is a projective variety, and this occurs just when \((X,G)\) is a flag variety, and just when \(P\) contains a maximal connected solvable subgroup \cite{Borel:1991} p. 148.
Every parabolic subgroup is connected p. 197 Proposition 14.18.

@@

\subsection{Opposite parabolic subgroups}

@@

(We can define such an automorphism explicitly as \(e_{\alpha}\mapsto -e_{-\alpha}\) on root vectors in a Chevalley basis; see section~\vref{subsubsection:ChevalleyBases}.)
Our automorphism sends \(P\) to its \emph{opposite} \(aP\) parabolic subgroup, so gives \(P\cap aP=\GZ\) and \(G_+\cap aG_+=\set{1}\) and an open subset of \(G\) consists of elements uniquely expressed as a product \(p,q\in P\times aG_+\mapsto pq\in G\) \cite{Cap/Slovak:2009} p. 294, \cite{Borel:1991} p. 198 Proposition 14.21.

@@

Clearly \(\breve{G}\) has Lie algebra containing all root vectors of all \(P\)-maximal and \(P\)-minimal roots, and is generated by the \(1\)-parameter subgroups of those root vectors.

@@

So \(\breve{X}=\breve{G}/\breve{P}=G'/P'\) is the Grassmannian of linear subspaces of dimension \(n_1\) inside \(V_1\oplus V_4\).

@@

If \((X,G)\) is almost effective then so is \((\breve{X},\breve{G})\).

@@

We will see \vpageref{subsection:Freedom} that the associated cominiscule subvariety of an irreducible flag variety \((X,G)\) satisfies an open condition on its tangent spaces, which we call \emph{freedom}.
We will see that the symmetry group of the associated cominiscule subvariety has strictly maximal dimension among symmetry groups of subvarieties of \(X\) with free smooth locus.
We will see that all free smooth subvarieties are homogeneous, evidence for the conjecture that the associated cominiscule subvariety is the unique free smooth subvariety.

@@

For a flag variety \(X=G/P\), the root system is also \emph{\(P\)-graded} by \(\sum n_i\), but summing only over the noncompact simple roots.

@@

Our hyperplane is always chosen so that we can see the heights, i.e. roots of a given height lie on a parallel hyperplane:

@@

Henceforth suppose that \(G\) is a connected complex semisimple Lie group.
\begin{lemma}\label{lemma:breve.G.connected}

@@

Every parabolic subgroup of a connected reductive Lie group is connected, so \(G\) and \(P\) are connected.

@@

The subgroup \(Z=Z_{G_+}\) is a Zariski closed subgroup of a unipotent linear algebraic group, so connected and isomorphic as an affine variety to complex Euclidean space \cite{Kambayashi/Miyanishi/Takeuchi:1974} p. 147 Remark A.3, \cite{Springer:2009} p. 243 corollary 14.2.7.

@@

expanding out into a sum with entirely positive coefficients unless \(\log g\) is a sum of \(P\)-maximal roots, so the center \(Z=Z_{G_+}\) of \(G_+\) consists precisely of the exponentials of elements of

@@

the sum being over the box, i.e. over the \(P\)-maximal roots, so \(Z\) is connected.
By definition of \(\breve{G}\), \(\breve{G}:=\left<Z,aZ\right>\) is connected, since \(Z\) is connected. 

@@

%\begin{lemma}
%Two associated cominiscule subvarieties are tangent at a point just when they are equal.
%\end{lemma}
%\begin{proof}
%Suppose that \(T_{x_0}\breve{X}=T_{x_0}g\breve{X}\).
%We can replace \(g\) by \(gh\) for any element \(h\in\breve{G}\), and \(\breve{G}\) acts transitively on \(\breve{X}\) by definition of \(\breve{X}\), so we can arrange that \(gx_0=x_0\), i.e. write \(g\) as \(p\) and we can assume that \(p\in P\).
%These tangent spaces are
%\[
%\breve\LieG/\breve\LieP=\Ad_p\left(\breve\LieG/\breve\LieP\right)\subseteq\LieG/\LieP.
%\]
%So
%\[
%\breve\LieG=\Ad_p\breve\LieG\subseteq\LieG.
%\]
%Hence taking exponentials, the identity components match:
%\[
%\breve{G}=\Ad_p\breve{G}.
%\]
%So their orbits are the same.
%\end{proof}

@@

It also suffices to assume that \((X,G)\) is an irreducible flag variety, as otherwise it is a product of irreducibles and the subgroup \(\breve{G}\) is the product of the associated subgroups, and so on.
The Lie algebra \(\LieZ\) of the center \(Z\) of the unipotent radical of \(P\) is the sum of the root spaces of the box, i.e. of the \(P\)-maximal roots.

@@

So \(a\LieZ\) is the sum of the root spaces of the roots of minimal \(P\)-grade, the opposite box.

@@

By lemma~\vref{lemma:breve.G.connected},  \(\breve{G}\) is a complex semisimple Lie group.
Since its root system lies inside that of \(G\), its Cartan subgroup is a subgroup of the Cartan subgroup of \(G\), generated by its coroots.
Note that \(P\) contains the Cartan subgroup of \(G\), hence that of \(\breve{G}\), and contains \(Z\), so contains the parabolic subgroup of \(\breve{G}\) generated by the box.
So this parabolic subgroup fixes the same point \(x_0\in X\) so lies in \(\breve{P}\).
The \(\breve{G}\)-orbit \(\breve{X}\) of that point \(x_0\) is a flag variety \((\breve{X},\breve{G})\), with stabilizer \(\breve{P}=P\cap\breve{G}\) a parabolic subgroup, so connected.

Since \(\breve\LieG\) is a complex subgroup of \(\LieG\), \(\breve{G}\) is a complex subgroup of \(G\), and so \(\breve{X}\subseteq X\) is a compact complex submanifold, and \(X\) is a smooth projective variety, so \(\breve{X}\subseteq X\) is a smooth subvariety.

@@

Since \(\LieZ\) is an irreducible \(\GZ\)-module, if we start at the highest root, say \(\aa_0\), we can pass from it via root strings to get to any \(P\)-maximal root, i.e. anywhere in the box, repeatedly passing between \(P\)-maximal roots \(\aa,\bb\) by subtracting a \(\breve{P}\)-compact positive root \(\aa-\bb\), so that bracketing a root vector \(\XX{\aa-\bb}\) of root \(\aa-\bb\) takes \(\XX{\aa}\) to a nonzero multiple of \(\XX{\bb}\).
Hence \(\LieZ\) is an irreducible \(\breveLieGZ\)-module, hence an irreducible \(\breve\LieP\)-module.
As we have seen~\vpageref{subsec:reducible}, the number of irreducible modules of the parabolic subgroup is the number of irreducible factors of the flag variety, so \(\breve{X}\) is an irreducible flag variety.

We next prove that \((\breve{X},\breve{G})\) is effective.
An element \(g \in \breve{G}\) acts trivially on \(\breve{X}\) just when \(gg_1\breve{P}=g_1\breve{P}\) for all \(g_1 \in \breve{G}\), i.e. just when \(g\) lies in all \(\breve{G}\)-conjugates of \(\breve{P}\).

The Weyl group acts transitively on bases of simple roots \cite{Serre:2001} p.33 theorem 2.
So we can reverse basis by conjugation by some element of \(G\).
Since \(\breve{P} \subset \breve{G}\) is a parabolic subgroup, conjugation on \(\breve{P}\) and \(P\) reverses the box and its opposite.
But \(\breve{G}_+ \cap a \breve{G}_+=1\) and \(a\breve{G}_0=\breve{G}_+\).
So an element \(g\in\breve{G}\) acting trivially on \(\breve{X}\) lies in \(\breve{G}_0\), the maximal reductive subgroup of \(\breve{P}\).
Acting trivially on \(T_{x_0}\breve{X}\), \(g\) acts trivially on \(\LieZ\).
Reversing, it acts trivially on \(\LieZ_-\), so acts trivially on \(\breve\LieG\), so lies in the center of \(\breve{G}\).
The center lies in the Cartan subgroup, hence in the Cartan subgroup of \(G\).
\end{proof}
\subsection{Automorphisms of the associated cominiscule}
\begin{example}
The flag variety \((X,G)=(B_2/P,B_2)\), where \(P\subseteq B_2\) is the Borel subgroup:
\[
\drawroots{B}{3} 
\]
has associated cominiscule
\[
\begin{tikzpicture}[baseline=-.5]
\begin{rootSystem}{B}
\roots
\parabolic{3}
\draw[/root system/grading,line width=.3cm] (square cs:x=1,y=1) -- (square cs:x=-1,y=-1);
\end{rootSystem}
\end{tikzpicture}
\]
a rational curve \((\breve{X},\breve{G})=(\Proj{1},\PSL{2})\), invariant under not only its automorphism group \(\breve{G}=A_1=\PSL{2}\), which we see in our diagram, but also, as we will see, under rescaling by this root:
\[
\begin{tikzpicture}[baseline=-.5]
\begin{rootSystem}{B}
\roots
\parabolic{3}
\draw[/root system/grading,line width=.3cm] (square cs:x=1,y=1) -- (square cs:x=-1,y=-1);
%\draw (square cs:x=-1,y=1) circle (2pt);
\draw[latex-,shorten <=1mm] (square cs:x=1,y=-1) --++ (.3cm,-.3cm);
\end{rootSystem}
\end{tikzpicture}
\]
The root is perpendicular to the \(\breve{G}\) root system, so commutes with the \(\breve{G}\) root vectors.
The vector field on \(X\) associated to the root vector of that root vanishes at our chosen point \(x_0\in X\) stabilized by \(P\), since the root lies in the Lie algebra of \(P\).
Since the vector field commutes with those of \(\breve{G}\), it is \(\breve{G}\)-invariant.
Since \(\breve{G}\) acts transitively on \(\breve{X}\), our vector field vanishes at all points of \(\breve{X}\).
Hence, in this example, the automorphism group \(G_{\breve{X}}\) of \(\breve{X}\) as a subvariety of \(X\) is larger than the automorphism group \(\breve{G}\) as a flag variety, containing the exponential of this root.

On the other hand, consider the root
\[
\begin{tikzpicture}[baseline=-.5]
\begin{rootSystem}{B}
\roots
\parabolic{3}
\draw[/root system/grading,line width=.3cm] (square cs:x=1,y=1) -- (square cs:x=-1,y=-1);
%\draw (square cs:x=1,y=-1) circle (2pt);
\draw[latex-,shorten <=1mm] (square cs:x=-1,y=1) --++ (-.3cm,.3cm);
\end{rootSystem}
\end{tikzpicture}
\]
It doesn't belong to the Lie algebra of \(P\), so doesn't vanish at \(x_0\).
Commuting with the root vectors of \(\breve{G}\), it is \(\breve{G}\)-invariant, so it doesn't vanish at any point of \(\breve{X}\).
If tangent to \(\breve{X}\) at some point of \(\breve{X}\), and commuting with the root vectors of \(\breve{G}\), it is tangent at every point, nowhere vanishing, not possible on the projective line \(\breve{X}=\Proj{1}\).
Hence this root vector is a vector field on \(X\), nowhere tangent to \(\breve{X}\).
\end{example}
\subsection{Computing the automorphism Lie algebra}
We return our thoughts to the general flag variety \((X,G)\).
By definition of \((\breve{X},\breve{G})\), \(\breve{X}\subset X\) is a smooth subvariety.
So the subgroup \(G'\subseteq G\) leaving \(\breve{X}\subseteq X\) invariant is a Zariski closed subgroup of \(G\), hence linear algebraic.
(As we have noted, we will find that, while \(G'\) contains \(\breve{G}\), it is often larger than \(\breve{G}\) and is not always semisimple.)
Let \(P':=G'\cap P\), so
\[
\breve{X}=\breve{G}/\breve{P}=G'/P'.
\]
Clearly \(\breve{X},\breve{P},\breve{G}\) are connected, so \(G'\) is connected just when \(P'\) is connected.
\begin{theorem}\label{theorem:aut.Lie.alg}
The automorphism group \(G'\) of the associated cominiscule subvariety \(\breve{X}\subset X\) of a flag variety is a complex linear algebraic group with Lie algebra precisely the sum of 
\begin{itemize}
\item
the \(P\)-maximal and \(P\)-minimal root spaces, 
\item
the maximal reductive \(\LieG_0\), and 
\item
the roots spaces of all positive roots perpendicular to all \(P\)-maximal roots.
\end{itemize}
\end{theorem}
\begin{proof}
Clearly \(G'\subseteq G\) is a complex linear algebraic subgroup containing \(\breve{G}\), hence acts transitively on \(\breve{X}\).
Let \(G'_0\subset G'\) be the connected component of the identity.
Since \(\breve{X}\) is connected and \(G'_0\) is a connected linear algebraic group acting transitively on \(\breve{X}\), the stabilizer \(P'_0:=P'\cap G'_0\) of the point \(x_0\in X\) is a parabolic subgroup of \(G'_0\) \cite{Borel:1991} p. 148, hence is connected.
Since \(\breve{X}=G'/P'\) is connected, \(P'\) intersects every component of \(G'\).

Since \(\breve{G}\) acts transitively on \(\breve{X}\), every element of \(G'\) is a product of an element of \(\breve{G}\) and an element of \(P\).
Clearly \(P'\) is the normalizer of \(\breve{G}P\) in \(P\).
Thus \(P'\) contains all elements of \(P\) which normalize \(\breve{G}\).
It contains \(\breve{P}\) and hence contains \(Z=Z_{G_+}\subseteq\breve{P}\).
The maximal reductive \(G_0\subseteq P\) normalizes \(P\) hence \(G_+\) hence \(Z\), and is invariant under the automorphism \(a\).
Thus \(G_0\subseteq P'\).
Hence \(P'\) contains the Cartan subgroup, so its Lie algebra is a sum of the Cartan subalgebra together with various root spaces.

By definition \(P'\) preserves \(\breve\LieG+\LieP\).
Since \(\breve{G},P\) are connected, so is \(\breve{G}P\), and so \(P'\) is precisely the subgroup of \(P\) preserving \(\breve\LieG+\LieP\subseteq\LieG\).
Take a \(P\)-submaximal \(P\)-positive root \(\aa\) at less than right angle with some \(P\)-maximal root \(\mx\bb\).
Then the root vectors of \(\aa\) and \(\mn\bb:=-\mx\bb\) bracket out of \(\breve\LieG+\LieP\).
Hence the Lie algebra \(\LieG'\) of \(G'\) contains no such.

The \(P\)-submaximal \(P\)-positive roots \(\aa\) at a right angle to all \(P\)-maximal roots \(\mx\bb\) have roots vectors commuting, so the root vectors of \(\aa\) preserve \(\breve\LieG\oplus\LieP\).
Roots which are \(P\)-positive are also positive, and conversely if they are not \(P\)-compact.
So \(\LieP'\) contains the root vectors of precisely the \(P\)-maximal roots and the \(P\)-positive roots perpendicular to them, and so consists precisely of these root vectors and the Cartan subalgebra.
